Particle Interception as a Function of Reynolds Number and Collector Density

### 2020-08-29 code/analysis.Rmd
analysis.pdf and analysis.rmd both on v2

## Data sourcing and tidying

### Changes from Version 1 (8/26)

@@

### Data background

@@

### Suspended sediment data

@@

### Sediment trap data

Data import and wrangling follows a similar structure to the pump data.

@@

### Velocity

We estimated Reynolds number from a flow velocity experiment in the open channel conducted in early 2019.

![](../pics/frequency_velocity_boxplots.jpeg)

It looks like there are a lot of outliers, but that's because there are simply a ton of Vectrino points and the distribution clearly has long tails. See how tiny the IQR boxes are, which by definition encompass 50% of points.

It should also be noted that the back-of-envelope estimate of $v = f / 500$ is extremely close, probably by design.

The data and code for this analysis is MIA at the moment, but should turn up soon.

### Flume volume

The test section is modeled as a rectangular prism, making calculation of volume trivial. However, the water in the flume is in an irregular form. In order to correct for the time water spends flowing outside the test section, where collectors are not acting on particles, we must estimate the total volume of water. We did so using a simple integration of the volumetric flow rate at the drain. We ran an experiment previously in 2018 or 2019, but my physics common sense lapsed: We picked the drain hose up off the ground to take measurements, then stuck it back down in the drain for most of the flow time. This led to an underestimate of about 5-10%. I re-did the analysis. 1st through 4th degree polynomial regression all yield the same volume +/- <0.5%:

```{r}
# flows in ml/s: 
flow <- c(131, 132, 132, 131, 134, 129, 130, 129, 130, 129, 130, 127, 127, 127, 126, 
          123, 126, 125, 124, 125, 127, 125, 123, 122, 125, 124, 118, 123, 125, 125, 
          120, 122, 124, 126, 122, 117, 123, 118, 120, 120, 121, 118, 118, 118, 119, 
          118, 120, 117, 117, 114, 115, 115, 116, 112, 112, 110, 111, 108, 105, 105, 
          102, 100, 100, 102)

# times in fractional hrs of day:  
time <- c(12.13333, 12.21667, 12.28333, 12.41667, 12.5, 12.56667, 12.6, 13.11667,
          13.15, 13.18333, 13.58333, 13.61667, 13.65, 14.25, 14.28333, 14.31667, 
          14.35, 14.38333, 14.68333, 14.71667, 14.75, 14.78333, 14.81667, 14.85, 
          14.88333, 14.93333, 15.35, 15.38333, 15.41667, 15.45, 15.48333, 15.76667, 
          15.8, 15.86667, 15.9, 15.93333, 15.96667, 16.18333, 16.23333, 16.28333, 
          16.31667, 16.33333, 16.36667, 16.55, 16.58333, 16.61667, 16.81667, 16.85, 
          16.9, 16.93333, 16.98333, 17.03333, 17.06667, 17.2, 17.23333, 17.35, 
          17.36667, 17.53333, 17.58333, 17.61667, 17.73333, 17.75, 17.78333, 17.8)

#flow start time:
t_0 <- 11.59166667

#time at which flume was drained to top of test-array holding basin
t_b <- 16.96666667

#time at which basin was drained
t_f <- 17.825

#polynomial model fits:
fits <- lapply(1:4, function(x) {lm(flow ~ poly(time, x, raw = TRUE))})

#finite integral calculator
poly_fin_integr <- function(model,lim1,lim2) {
  sum(coefficients(model)/(1:length(coefficients(model)))*
        lim2^(1:length(coefficients(model))))-
  sum(coefficients(model)/(1:length(coefficients(model)))*
        lim1^(1:length(coefficients(model))))
}

estimates <- lapply(fits, function(x) poly_fin_integr(x, lim1 = t_0, lim2 = t_b))

#convert from mL/s*h to cubic m: 1 mL/s*h = 3600 mL = 3.6 L = 0.0036 cubic m

estimates <- unlist(estimates)*.0036

names(estimates) <- c("linear", "quadratic", "cubic", "4th order")

estimates

@@

## Modelling for eta (and all the other vars of interest)

@@

Here is a nice function found on StackExchange.com, which performs error propogation in R using derivatives:

@@

We estimate k_tot from our pump data by linear regression of `log(mvc)`:

```{r, warning=FALSE}

@@

final %>%
  ggplot(aes(x = as.numeric(pump_freq), y = k_c, color = dowel_density)) +
  geom_line()

>>>>

## Data Tidying

### CHANGES FROM VERSION 1 (8/26)

@@

### Time-decay runs

@@

@@

For the sediment traps, data import and wrangling follows a similar structure to the pump data.

@@

### Velocity experiment

We estimated Reynolds number from a flow velocity experiment in the open channel conducted on Jan 28, 2019. The Vectrino data is in commit `e78357` and the R code for the regression in commit `17c8af` in the esdlflume Github repo (Thanks Candace!).

There's an image in version 1 (and in /pics) but the linear regression is simply $u = 0.00081282 + 0.00195723f$, where $u$ is flow velocity (m/s) and $f$ is pump frequency (Hz). Hence our velocities at 10, 20, and 30 Hz, were `r 0.00081282 + 0.00195723*10`, `r 0.00081282 + 0.00195723*20`, and `r 0.00081282 + 0.00195723*30` respectively.

It should be noted that the back-of-envelope estimate of $v = f / 500$ is very close for our range, probably by design. For the regression, $R^2 > .99$.

### Flume volume experiment

The test section is modeled as a rectangular prism, making calculation of volume trivial. However, the water in the flume is in an irregular form. In order to correct for the time water spends flowing outside the test section, where collectors are not acting on particles, we must estimate the total volume of water. We did so using a simple integration of the volumetric flow rate at the drain. We ran an experiment previously in 2018 or 2019, but my physics common sense lapsed: We picked the drain hose up off the ground to take measurements, then stuck it back down in the drain for most of the flow time. This led to an underestimate of about 5-10%. I re-did the analysis. 1st through 4th degree polynomial regression all yield the same volume +/- <0.5% (code in this repo, under /code)

@@

## Modelling for ECE%

@@

Here is a function found on StackExchange.com, which performs error propogation in R using derivatives:

@@

We estimate `k_tot` from our pump data by linear regression of `log(mvc)`:

```{r, warning=FALSE, message=FALSE}

@@

final %>%
  filter(dowel_density != "0000") %>%
  ggplot(aes(x = as.numeric(pump_freq), y = k_c, color = dowel_density)) +
  geom_line() +
  geom_point()
